A thesis draft

### 2018-11-25 res/sections/15-ArchitecturalImplementation.tex
Start second attempt part

After the initial Openstack and Openbaton configuration we started making the
components talking to each other. First, we tried a simple \verb!iperf! data
transmission with two VNFs deployed by Openbaton as a common hypervisor-based
Openstack instances, as explained in~\cite{openbatonIperf}. With this test
completed, we started tinkering Openstack for using Docker containers instead of
traditional hypervisor instances. When we started developing this solution, we
adopted the Zun module, that aims to smoothly integrate Docker in Openstack.
Unfortunately, we were not able to make it properly work, in particular, it was
not possible to open selected desired ports on the Docker containers. This lead
to the impossibility to perform packet forwarding between VNFs. On top of that,
we found difficulties regarding the integration of Docker images in TOSCA
definitions and Openstack. These two issues are related: the TOSCA definition
does not have any keyword to specificate in any way that the image used for the
VNF definition is a Docker image, and Openstack, with the Zun plugin, at the
time of the implementation did not performed the image pull from any Docker
registry. The impossibility to easily set the containers port and the lack of
some support in between the TOSCA definition and the Openstack Zun plugin pushed
us to drop Openstack as a possible solution and to start looking for another
valid VIM component.

>>>>

After the initial Openstack and Openbaton configuration we started making the
components talking to each other. First, we tried a simple \verb!iperf! data
transmission with two VNFs deployed by Openbaton as a common hypervisor-based
Openstack instances, as explained in~\cite{openbatonIperf}. With this test
completed, we started tinkering Openstack for using Docker containers instead of
traditional hypervisor instances. When we started developing this solution, we
adopted the Zun module, that aims to smoothly integrate Docker in Openstack.
Unfortunately, we were not able to make it properly work, in particular, it was
not possible to open desired ports on the Docker containers. This lead to the
impossibility to perform packet forwarding between VNFs. On top of that, we
found difficulties regarding the integration of Docker images in TOSCA
definitions and Openstack. These two issues are related: the TOSCA definition
does not have any keyword to specificate in any way that the image used for the
VNF definition is a Docker image, and Openstack, with the Zun plugin, at the
time of the implementation did not performed the image pull from any Docker
registry. The impossibility to easily set the containers ports and the lack of
some support in between the TOSCA definition and the Openstack Zun plugin pushed
us to drop Openstack as a possible solution and to start looking for another
valid VIM component.

\section{Second attempt}

After the failure to find a viable setting for Openstack, we decided to look
around for other solutions. We started studing proper Docker container
orchestrators, that could offer all the features we needed: container lifecycle
management, scaling, healing and monitoring. We decided to pick two of the most
mature and famous ones: Docker Swarm and Kubenetes. The decision to study only
these two relies on their big community supporting them and the possibility to
easily found projects that used the forementioned frameworks.

\subsection{Docker Swarm}

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.35]{swarm_diagram}
  \caption[Docker Swarm schema]{Docker Swarm schema~\cite{dockerSwarmWiki}.}
  \label{chap:archimpl:sec:secondattempt:img:dockerswarm}
\end{figure}

First we started playing with Docker Swarm. Docker Swarm is a product directly
made by the Docker Inc., so it seamlessly integrates with other Docker tools
such as Docker Compose. The installation and set-up is pretty straightforward,
and the configuration is almost automatically handed by Docker itself. The
minimum number of nodes in order to create a proper Docker Swarm cluster is of
free nodes, where one must be a manager node (designed to orchestrate the other
nodes in the cluster), while the other can be simple workers. Multiple managers
can coexist in the same cluster, as long as only one effectively operates. The
others act like backup manager in case the one orchestrating goes down for some
reason. The networking is given out-of-the box, and the same holds for
distributed persistence. While this for most of the uses cases can be considered
a pro, in our specific situation the impossibility to tweak the network
configuration was a cons, since it was important for us to have the full control
over the network, especially regarding the SFC implementation, explained better
in Chapter~\ref{chap:vnf_ns_impl}.

\subsection{Kubernetes}
